Rethinking how students learn to design programs

### 2016-04-23 manuscript-body.tex
Gut code samples; replace with figures

After declaring variables and opening the three provided text files (flights, routes, and airports)\footnote{See Appendix A for examples of each of the three file types.}, Rebecca's one-stop flight code enters a series of conditionally-nested \texttt{fscanf()} commands. The purpose of \texttt{fscanf()} is to scan through the characters of a file, most often scanning one line at a time. The programmer can specify delimited patterns of text to look for, e.g., ``in each line, look for a word, followed by a space, followed by two digits.'' \texttt{fscanf()} also gives the programmer the flexibility to store the values of matched patterns to variables. Here's an annotated explanation of file-scanning pattern Rebecca uses on line 24 of her code:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fscanf(               }\CommentTok{// Read in a line of data}
  \NormalTok{routes,             }\CommentTok{// From the file called 'routes'}
  \StringTok{"%d.  %s %s"}\NormalTok{,       }\CommentTok{// Match the pattern (integer) (string) (string)}
  \NormalTok{&route_id,          }\CommentTok{// Store the value of integer to the variable route_id}
  \NormalTok{departure_airport,  }\CommentTok{// Store the value of the first string to departure_airport}
  \NormalTok{arrival_airport     }\CommentTok{// Store the value of the second string to arrival_airport}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

``Once you find a word followed by two digits, store the word in the following location in memory, and store the digits in this other location in memory.'' What's interesting isn't \emph{that} Rebecca used \texttt{fscanf()} to read through files like the list of airports. Any suitable solution for this project would need extract text patterns from source files, which \texttt{fscanf()} does. Rather, what's interesting is \emph{how} she uses \texttt{fscanf()} in her design.

@@

\hypertarget{fscanfPattern}{\label{fscanfPattern}}
\begin{Shaded}
\begin{Highlighting}[numbers=left,,firstnumber=20,]
\KeywordTok{while}\NormalTok{(fscanf(airports, }\StringTok{"%s  %50[^}\CharTok{\textbackslash{}n}\StringTok{]"}\NormalTok{, airport_code, city) != EOF)   }\CommentTok{//scan for the departure code but not to the arrival code}
  \NormalTok{\{}
    \KeywordTok{if}\NormalTok{(airport_code != departure_input || airport_code != arrival_input)}
      \NormalTok{\{}
        \KeywordTok{while}\NormalTok{(fscanf(routes, }\StringTok{"%d.  %s %s"}\NormalTok{, &route_id, departure_airport, arrival_airport) != EOF)  }\CommentTok{//finds the name of the city of the departure airport}
          \NormalTok{\{}
            \KeywordTok{if}\NormalTok{(departure_input[}\DecValTok{0}\NormalTok{]==departure_airport[}\DecValTok{0}\NormalTok{] && departure_input[}\DecValTok{1}\NormalTok{]==departure_airport[}\DecValTok{1}\NormalTok{] && departure_input[}\DecValTok{2}\NormalTok{]==departure_airport[}\DecValTok{2}\NormalTok{])}
              \NormalTok{\{}
                \KeywordTok{if}\NormalTok{(arrival_airport[}\DecValTok{0}\NormalTok{] != arrival_input[}\DecValTok{0}\NormalTok{] || arrival_airport[}\DecValTok{1}\NormalTok{] != arrival_input[}\DecValTok{1}\NormalTok{] || arrival_airport[}\DecValTok{2}\NormalTok{] != arrival_input[}\DecValTok{2}\NormalTok{])}
                  \NormalTok{\{}
                    \NormalTok{strcpy(stop_code, arrival_airport);}
                    \NormalTok{rewind(routes);}
                    \NormalTok{departure_airport[}\DecValTok{0}\NormalTok{]=' ';}
                    \NormalTok{departure_airport[}\DecValTok{1}\NormalTok{]=' ';}
                    \NormalTok{departure_airport[}\DecValTok{2}\NormalTok{]=' ';}
                    \NormalTok{arrival_airport[}\DecValTok{0}\NormalTok{]=' ';}
                    \NormalTok{arrival_airport[}\DecValTok{1}\NormalTok{]=' ';}
                    \NormalTok{arrival_airport[}\DecValTok{2}\NormalTok{]=' ';}
                    \KeywordTok{while}\NormalTok{(fscanf(routes, }\StringTok{"%d.  %s %s"}\NormalTok{, &route_id, departure_airport, arrival_airport) != EOF)  }\CommentTok{//now searches for the place where the stop is the}
                      \NormalTok{\{                                                                                      }\CommentTok{//depart airport and the arrival input is correct}
\end{Highlighting}
\end{Shaded}

>>>>

After declaring variables and opening the three provided text files (flights, routes, and airports)\footnote{See Appendix A for examples of each of the three file types.}, Rebecca's one-stop flight code enters a series of conditionally-nested \texttt{fscanf()} commands. Any suitable solution for this project would need extract text patterns from source files, which \texttt{fscanf()} does. So, \emph{that} Rebecca uses \texttt{fscanf()} in her code is not itself surprising. Rather, what's interesting is \emph{how} she uses \texttt{fscanf()} in her design.

@@

\begin{figure}[htbp]
\centering
\includegraphics{media/fscanf.pdf}
\caption{A portion of Rebecca's nested \texttt{fscanf()} statements}
\end{figure}
